EGGs are adhesive!\\ An algebraic account of equality saturation

### 2025-08-23 jlamp/jlamp.tex
minima

\begin{defi} Let $\mathcal{G}$ be an hypergraph. A node $v\in V_{\mathcal{G}}$ is \emph{isolated} if it does not occur as a letter in either $s_{\mathcal{G}}(e)$ or in
$t_{\mathcal{G}}(e)$ for every $e\in E_{\mathcal{G}}$. $\mathcal{G}$ is \emph{pruned} if it has no isolated nodes. We denote by $\phyp$ and $\pshyp$ the full subcategories of $\hyp$ given by, respectively, pruned and pruned and separated hypergraphs.
\end{defi}

@@

\begin{rem}\label{rem:mprun}
	Let $(f,g)$ and $(h, k)$ be two morphisms $\mathcal{G}\rightrightarrows \mathcal{H}$ of $\phyp$ and suppose that $f=h$. We can use the absence of isolated nodes to show that $g=k$ too. Let $v\in V_{\mathcal{G}}$, we have two cases:
	\begin{itemize}
		\item $v$ occurs in $s_{\mathcal{G}}(e)$ for some $e\in E_{\mathcal{G}}$. Let $1\leq i\leq \lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ be such that $v=s_{\mathcal{G}}(e)_i$, then we have
		\[g(v)=g(s_{\mathcal{G}}(e)_i)=(g^\star(s_{\mathcal{G}}(e)))_i=s_{\mathcal{H}}(f(e))_i=s_{\mathcal{H}}(h(e))_i=(k^\star(s_{\mathcal{G}}(e)))_i=k(s_{\mathcal{G}}(e)_i)=k(v)\]
		\item $v$ occurs in $t_{\mathcal{H}}(e)$ for some $e\in E_{\mathcal{H}}$. Let $1\leq i\leq \lgh_{V_{\mathcal{H}}}(t_{\mathcal{G}}(e))$ be such that $v=t_{\mathcal{G}}(e)_i$, then we have
		\[g(v)=g(t_{\mathcal{G}}(e)_i)=(g^\star(t_{\mathcal{G}}(e)))_i=t_{\mathcal{H}}(f(e))_i=t_{\mathcal{H}}(h(e))_i=(k^\star(t_{\mathcal{G}}(e)))_i=k(t_{\mathcal{G}}(e)_i)=k(v)\]
	\end{itemize}
	In both cases $\hat{g}(v)=k(v)$ and so $\hat{g}= k$ as wanted.
\end{rem}

@@

If we restrict to pruned separated hypergraph the previous issue disappeares.  The crucial observation is that the a pruned separated hypergraph is characterized by its set of hyperedges and the length of their sources and targets. See \Cref{proof:pruned} for a proof.

@@

\begin{defi}
	Given an hypergraph $\mathcal{G}$, the category $\hyp_\mathcal{G}$ of hypergraphs \emph{labelled by $\mathcal{G}$}  or of \emph{$\mathcal{G}$-hypergraphs} is $\hyp/\mathcal{G}$.
\end{defi}

@@

\begin{exa}\label[example]{lab_2}
	Consider again $\Sigma$ the signature of \Cref{exa_3}, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled by a morphism
	$(l, !_{V_\mathcal{G}}): \mathcal{G} \to \mathcal{G}^\Sigma$ that is characterised by the image of the edges. 
	If $l(h_1) = a$, $l(h_2) = 2$, $l(h_3) = *$, and $l(h_4) = /$, we represent it visually 
	by putting the labels of the edges in  $\mathcal{G}^\Sigma$ inside the boxes representing the edges of $\mathcal{G}$.
		\begin{center}
		\begin{tikzpicture}
			\begin{pgfonlayer}{nodelayer}
				\node[style=none] (lab1) at (-1.8, 0.8) {};
				\node[style=small box] (h1) at (-1.8, 0.3) {$a$};
				\node[style=small box] (2) at (-1.8, -0.5) {$2$};
				\node[style=none] (lab2) at (-1.8, -1.0) {};
				\node[style=none] (h1b) at (-1.8, 0.3) {};
				\node[style=none] (th2) at (0, 0.3) {};
				\node[style=none] (bh2) at (0, -0.3) {};
				\node[style=none] (h2b) at (0, 0) {};
				\node[style=none] (th3) at (1.8, 0){};
				\node[style=none] (bh3) at (1.8, -0.6){};
				\node[style=none] (h3c) at (1.8, -0.3){};
				\node[style=none] (2c) at (-1.8, -0.5){};
				\node[style=none] (lab3) at (0, 0.8) {};
				\node[style=medium box] (h2) at (0, 0) {$*$};
				\node[style=none] (lab4) at (1.8, 0.5) {};
				\node[style=medium box] (h3) at (1.8, -0.3) {$/$};
				\node[style=none](labv2) at (-0.9, -0.9) {};
				\node[style=node] (v2) at (-0.9, -0.5) {};
				\node[style=none] (labv1) at (-0.9, 0.7) {};
				\node[style=node] (v1) at (-0.9, 0.3) {};
				\node[style=none] (labv3) at (0.9, 0.4) {};
				\node[style=node] (v3) at (0.9, 0) {};
				\node[style=none] (labv4) at (2.7, 0.1) {};
				\node[style=node] (v4) at (2.7, -0.3) {};
			\end{pgfonlayer}
			\begin{pgfonlayer}{edgelayer}
				\draw (h1b.center) to (th2.center);
				\draw (2c.center) to (v2);
				\draw[in=180, out=30] (v2) to (bh2.center);
				\draw (h2b.center) to (th3.center);
				\draw (v4) to (h3c.center);
				%\draw (h3c) to (v4);
				\draw[in=180, out=-30](v2) to (bh3);
			\end{pgfonlayer}
        	\end{tikzpicture}
	\end{center}
%
	\iffalse
	\begin{center}\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_1$}] (A) at (0,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_2$}] (B) at (0,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_3$}] (C) at (3,-0.75) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_4$}] (D) at (3,-2.25) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=right:{$v_5$}] (E) at (6,-1.5) {};
			\draw[->-=.5] (1.75,-0.75)--(C)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (1.25, -1) rectangle (1.75, -0.5) {};
			\draw[->-=.5] (4.75,-1.5)--(E)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (-1.5,0)--(A)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (4.25, -1.75) rectangle (4.75, -1.25) {};
			\node at (4.5, -1.05){$h_3$};
			\node at (1.5, -0.3){$h_2$};
			\node at (-1.75, 0.45){$h_1$};
			\draw[rounded corners] (-2, -0.25) rectangle (-1.5, 0.25) {};
			\draw[->-=.5] (A)..controls(0.5,0)and(1.2,-0.2)..(1.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (B)..controls(0.5,-1.5)and(1.2,-1.3)..(1.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

@@

\begin{thm} Let $\mathcal{G}$ be an hypergraph, then %the following fact hold
%	\begin{enumerate}
%		\item an arrow $(f,g)$  if and only if is so in $\lmo$; \marginpar{ovvero?}
		%\item 
		$\lmo_{\mathcal{G}}$ is quasiadhesive.
%	\end{enumerate}
\end{thm}

@@

\dx*
\begin{proof}\label{proof:dx}
	Let $\mathcal{G}$ be an hypergraph. For every $e\in E_{\mathcal{G}}$, let $n_e=\lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ and $m_e=\lgh_{V_{\mathcal{G}}}(t_{\mathcal{G}}(e))$. Given $1\leq i\leq n_e+m_e$,  we can define $\delta^e_i\colon \mto V_{\mathcal{G}}$ as
	\[\delta^e_i(0) =\begin{cases}
		s_{\mathcal{G}}(e)_i & i\leq n_e\\
		t_{\mathcal{G}}(e)_{i-n_e} &  n_e < i
	\end{cases}\]

@@

	\smallskip 
	Hence $(\id{E_{\mathcal{G}}}, \epsilon_{G})$ is a morphism $P(\sp(\mathcal{G}))\to \mathcal{G}$. Now, let $(f,g)\colon P(\mathcal{H})\to \mathcal{G}$ be a  morphism of pruned hypergraphs with $\mathcal{H}$ separated. Given a node $v$ in $V_{\mathcal{H}}$ we know that there exists exactly one $e_v$ such that $v$ occurs in $s_{\mathcal{H}}(e_v)$ or in $t_{\mathcal{H}}(e_v)$. Moreover, $v$ occurs exactly once and not in both. Let $i_v$ be the index at which it appears. We define
	\[\hat{g}(v):=\begin{cases}
		\iota_{f(e_v)}(\iota^{f(e_v)}_{i_v}(0)) & v \text{ occurs in } s_{\mathcal{H}}(e)\\
		\iota_{f(e_v)}(\iota^{f(e_v)}_{n_e + i_v}(0)) & v \text{ occurs in } t_{\mathcal{H}}(e)\\
	\end{cases}\]
	Computing we get
	\begin{gather*}
		\hat{g}^\star(s_{\mathcal{H}}(e))= \hat{g}^\star(\prod_{i=1}^{n_e} s_{\mathcal{H}}(e)_i)=\prod_{i=1}^{n_e}(\hat{g}(s_{\mathcal{H}}(e)_i))=\prod_{i=1}^{n_e}\iota_{f(e)}( \iota^{f(e)}_i(0))=s_{\sp(\mathcal{G})}(f(e))\\
		\hat{g}^\star(t_{\mathcal{H}}(e))= \hat{g}^\star(\prod_{i=1}^{m_e} t_{\mathcal{H}}(e)_i)=\prod_{i=1}^{m_e}(\hat{g}(t_{\mathcal{H}}(e)_i))=\prod_{i=1}^{m_e}\iota_{f(e)}( \iota^{f(e)}_{n_e+i}(0))=\prod_{i=n_e + 1}^{n_e + m_e}\iota_{f(e)}( \iota^{f(e)}_{i}(0))=t_{\sp(\mathcal{G})}(f(e))
	\end{gather*}
	proving that $(f, \hat{g})\colon \mathcal{H}\to\sp( \mathcal{G})$ is a morphism of $\pshyp$. Notice that	
	\begin{align*}
		&\epsilon_\mathcal{G}(\hat{g}(v))=\begin{cases}
			\epsilon_\mathcal{G}(\iota_{f(e_v)}(\iota^{f(e_v)}_{i_v}(0))) & v \text{ occurs in } s_{\mathcal{H}}(e_v)\\
			\epsilon_\mathcal{G}(\iota_{f(e_v)}(\iota^{f(e_v)}_{n_e + i_v}(0))) & v \text{ occurs in } t_{\mathcal{H}}(e_v)
		\end{cases}=	\begin{cases}
		\delta^{f(e_v)}_{i_v}(0)& v \text{ occurs in } s_{\mathcal{G}}(e)\\
		\delta^{f(e_v)}_{i_v}(0)& v \text{ occurs in } t_{\mathcal{G}}(e)
		\end{cases}\\ =&
		\begin{cases}
			s_{\mathcal{G}}(f(e_v))_{i_v}& v \text{ occurs in } s_{\mathcal{G}}(e)\\
			t_{\mathcal{G}}(f(e_v))_{i_v}& v \text{ occurs in } t_{\mathcal{G}}(e)
		\end{cases}= \begin{cases}
		(g^\star(s_{\mathcal{H}}(e_v)))_{i_v}& v \text{ occurs in } s_{\mathcal{G}}(e)\\
		(g^\star(t_{\mathcal{H}}(e_v)))_{i_v}& v \text{ occurs in } t_{\mathcal{G}}(e)\\
		\end{cases}\\=&\begin{cases}
		g(s_{\mathcal{H}}(e_v)_{i_v})& v \text{ occurs in } s_{\mathcal{G}}(e)\\
		g(t_{\mathcal{H}}(e_v)_{i_v})& v \text{ occurs in } t_{\mathcal{G}}(e)\\
		\end{cases}=g(v)
	\end{align*}
and so $(\id{E_{\mathcal{G}}}, \epsilon_{G})\circ P(f, \hat{g})=(f,g)$.

@@

\begin{proof}{\label{proof:pruned}}
	Let $\mathcal{G}$ be a pruned and separated hypergraph. We can define $F(\mathcal{G})$ as $(\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}}) \colon E_{\mathcal{G}}\to \mathbb{N}\times \mathbb{N}$. Given a morphism $(h,k)\colon \mathcal{G}\to \mathcal{H}$ and using \Cref{rem:le} we	have
	\[(\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}})\circ h = (\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}\circ h, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}}\circ h)=(\lgh_{V_{\mathcal{H}}}\circ k^\star\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{H}}}\circ k^\star\circ t_{\mathcal{G}})= (\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}})\]
	Thus we can define $F(h,k)$ simply as $h$ to get a functor $\pshyp\to \Set/(\mathbb{N}\times \mathbb{N})$. Let us show that it is an equivalence.
	\begin{itemize}
		\item $F$ is faithful. This follows at once from \Cref{rem:prun}.
		\item $F$ is full. Let $k\colon E_{\mathcal{G}}\to E_{\mathcal{H}}$ be an arrow such that 
		\[(\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}})\circ f =(\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}})\]

@@

		\noindent
		\begin{minipage}[l]{.78\linewidth}
		Since $\mathcal{G}$ is pruned and separated, $(V_{\mathcal{G}}, \{\iota^s_e\}_{e\in E_{\mathcal{G}}}+\{\iota^t_e\}_{e\in E_{\mathcal{G}}})$ is a coproduct. Given $v\in V^s_e$ and $w\in V^t_e$, let $i_v$ and $k_w$ be the unique indexes such that $\iota^s_e(v)=s_{\mathcal{G}}(e)_{i_v}$ and $\iota^t_e(w)= t_{\mathcal{G}}(e)_{k_w}$ and define:
			\[\begin{split}
				g^s_e\colon V^s_{e} &\to V_{\mathcal{H}}\\
				v&\mapsto  s_{\mathcal{H}}(f(e))_{i_v}
			\end{split} \qquad \begin{split}
				g^t_e\colon V^t_{e} &\to V_{\mathcal{H}}\\
				w&\mapsto  t_{\mathcal{H}}(f(e))_{k_w}
			\end{split}\]
			so that there is an arrow $g\colon V_ {\mathcal{G}}\to \mathcal{H}$ fitting in the diagram aside.
		\end{minipage}\hfill
		\begin{minipage}[r]{.2\linewidth}
		\xymatrix@R=14pt@C=14pt{&V^s_{\mathcal{G}} \ar[dr]^{g^s_e} \ar@{>->}[dl]_{\iota^s_e}\\ V_{\mathcal{G}} \ar@{.>}[rr]^-{g} && V_{\mathcal{H}}}		
		\xymatrix@R=14pt@C=14pt{&V^t_{\mathcal{G}} \ar[dr]^{g^t_e} \ar@{>->}[dl]_{\iota^t_e}\\ V_{\mathcal{G}} \ar@{.>}[rr]^-{g} && V_{\mathcal{H}}}		
		\end{minipage}

		Given $e\in E_\mathcal{G}$, for every $1\leq i \leq \lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ and $1\leq k \leq \lgh_{V_{\mathcal{G}}}(t_{\mathcal{G}}(e))$ we can compute to get:
		\begin{align*}
		(g^\star(s_{\mathcal{G}}(e)))_i&=(g^\star(s_{\mathcal{G}}(e)))_i=g(s_{\mathcal{G}}(e)_i)=s_{\mathcal{H}}(f(e))_i\\
		(g^\star(t_{\mathcal{G}}(e)))_k&=(g^\star(t_{\mathcal{G}}(e)))_k=g(t_{\mathcal{G}}(e)_k)=t_{\mathcal{H}}(f(e))_k
		\end{align*}

>>>>

\begin{defi} Let $\mathcal{G}$ be an hypergraph. A node $v\in V_{\mathcal{G}}$ is \emph{isolated} if it does not occur as a letter in either $s_{\mathcal{G}}(e)$ or in
$t_{\mathcal{G}}(e)$ for any $e\in E_{\mathcal{G}}$. $\mathcal{G}$ is \emph{pruned} if it has no isolated nodes. We denote by $\phyp$ and $\pshyp$ the full subcategories of $\hyp$ given by, respectively, pruned and pruned and separated hypergraphs.
\end{defi}

@@

\begin{rem}\label{rem:mprun}
	Let $(f,g)$ and $(h, k)$ be two morphisms $\mathcal{G}\rightrightarrows \mathcal{H}$ of $\phyp$ and suppose that $f=h$. We can use the absence of isolated nodes to show that $g=k$ too. Let $v\in V_{\mathcal{G}}$: we have two cases
	\begin{itemize}
		\item $v$ occurs in $s_{\mathcal{G}}(e)$ for some $e\in E_{\mathcal{G}}$. Let $1\leq i\leq \lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ be such that $v=s_{\mathcal{G}}(e)_i$, then we have
		\[g(v)=g(s_{\mathcal{G}}(e)_i)=(g^\star(s_{\mathcal{G}}(e)))_i=s_{\mathcal{H}}(f(e))_i=s_{\mathcal{H}}(h(e))_i=(k^\star(s_{\mathcal{G}}(e)))_i=k(s_{\mathcal{G}}(e)_i)=k(v)\]
		\item $v$ occurs in $t_{\mathcal{H}}(e)$ for some $e\in E_{\mathcal{H}}$. Let $1\leq i\leq \lgh_{V_{\mathcal{H}}}(t_{\mathcal{G}}(e))$ be such that $v=t_{\mathcal{G}}(e)_i$, then we have
		\[g(v)=g(t_{\mathcal{G}}(e)_i)=(g^\star(t_{\mathcal{G}}(e)))_i=t_{\mathcal{H}}(f(e))_i=t_{\mathcal{H}}(h(e))_i=(k^\star(t_{\mathcal{G}}(e)))_i=k(t_{\mathcal{G}}(e)_i)=k(v)\]
	\end{itemize}
	In both cases $\hat{g}(v)=k(v)$ and so $\hat{g}= k$ as wanted.
\end{rem}

@@

\smallskip
The previous issue disappears if we restrict to pruned separated hypergraphs.  The crucial observation is that the a pruned separated hypergraph is characterized by its set of hyperedges and the length of their sources and targets. See \Cref{proof:pruned} for a proof.

@@

\begin{defi}
	Let $\mathcal{G}$ be an hypergraph. The category $\hyp_\mathcal{G}$ of hypergraphs \emph{labelled by $\mathcal{G}$}  or of \emph{$\mathcal{G}$-hypergraphs} is $\hyp/\mathcal{G}$.
\end{defi}

@@

\begin{exa}\label[example]{lab_2}
	Consider again $\Sigma$ the signature of \Cref{exa_3}, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled by a morphism
	$(l, !_{V_\mathcal{G}}): \mathcal{G} \to \mathcal{G}^\Sigma$ that is characterised by the image of the edges. 
	If $l(h_1) = a$, $l(h_2) = 2$, $l(h_3) = *$, and $l(h_4) = /$, we represent it visually 
	by putting the labels of the edges in  $\mathcal{G}^\Sigma$ inside the boxes representing the edges of $\mathcal{G}$
		\begin{center}
		\begin{tikzpicture}
			\begin{pgfonlayer}{nodelayer}
				\node[style=none] (lab1) at (-1.8, 0.8) {};
				\node[style=small box] (h1) at (-1.8, 0.3) {$a$};
				\node[style=small box] (2) at (-1.8, -0.5) {$2$};
				\node[style=none] (lab2) at (-1.8, -1.0) {};
				\node[style=none] (h1b) at (-1.8, 0.3) {};
				\node[style=none] (th2) at (0, 0.3) {};
				\node[style=none] (bh2) at (0, -0.3) {};
				\node[style=none] (h2b) at (0, 0) {};
				\node[style=none] (th3) at (1.8, 0){};
				\node[style=none] (bh3) at (1.8, -0.6){};
				\node[style=none] (h3c) at (1.8, -0.3){};
				\node[style=none] (2c) at (-1.8, -0.5){};
				\node[style=none] (lab3) at (0, 0.8) {};
				\node[style=medium box] (h2) at (0, 0) {$*$};
				\node[style=none] (lab4) at (1.8, 0.5) {};
				\node[style=medium box] (h3) at (1.8, -0.3) {$/$};
				\node[style=none](labv2) at (-0.9, -0.9) {};
				\node[style=node] (v2) at (-0.9, -0.5) {};
				\node[style=none] (labv1) at (-0.9, 0.7) {};
				\node[style=node] (v1) at (-0.9, 0.3) {};
				\node[style=none] (labv3) at (0.9, 0.4) {};
				\node[style=node] (v3) at (0.9, 0) {};
				\node[style=none] (labv4) at (2.7, 0.1) {};
				\node[style=node] (v4) at (2.7, -0.3) {};
			\end{pgfonlayer}
			\begin{pgfonlayer}{edgelayer}
				\draw (h1b.center) to (th2.center);
				\draw (2c.center) to (v2);
				\draw[in=180, out=30] (v2) to (bh2.center);
				\draw (h2b.center) to (th3.center);
				\draw (v4) to (h3c.center);
				%\draw (h3c) to (v4);
				\draw[in=180, out=-30](v2) to (bh3);
			\end{pgfonlayer}
        	\end{tikzpicture}
	\end{center}
%
	\iffalse
	\begin{center}\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_1$}] (A) at (0,0) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_2$}] (B) at (0,-1.5) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_3$}] (C) at (3,-0.75) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v_4$}] (D) at (3,-2.25) {};
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=right:{$v_5$}] (E) at (6,-1.5) {};
			\draw[->-=.5] (1.75,-0.75)--(C)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (1.25, -1) rectangle (1.75, -0.5) {};
			\draw[->-=.5] (4.75,-1.5)--(E)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (-1.5,0)--(A)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[rounded corners] (4.25, -1.75) rectangle (4.75, -1.25) {};
			\node at (4.5, -1.05){$h_3$};
			\node at (1.5, -0.3){$h_2$};
			\node at (-1.75, 0.45){$h_1$};
			\draw[rounded corners] (-2, -0.25) rectangle (-1.5, 0.25) {};
			\draw[->-=.5] (A)..controls(0.5,0)and(1.2,-0.2)..(1.25,-0.6)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (B)..controls(0.5,-1.5)and(1.2,-1.3)..(1.25,-0.9)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};

@@

\begin{thm} Let $\mathcal{G}$ be an hypergraph. Then %the following fact hold
%	\begin{enumerate}
%		\item an arrow $(f,g)$  if and only if is so in $\lmo$; \marginpar{ovvero?}
		%\item 
		$\lmo_{\mathcal{G}}$ is quasiadhesive.
%	\end{enumerate}
\end{thm}

@@

\dx*
\begin{proof}\label{proof:dx}
	Let $\mathcal{G}$ be an hypergraph. For every $e\in E_{\mathcal{G}}$, let $n_e=\lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ and $m_e=\lgh_{V_{\mathcal{G}}}(t_{\mathcal{G}}(e))$. Given $1\leq i\leq n_e+m_e$,  we can define $\delta^e_i\colon \mto V_{\mathcal{G}}$ as
	\[\delta^e_i(0) =\begin{cases}
		s_{\mathcal{G}}(e)_i & \mbox{if } i\leq n_e\\
		t_{\mathcal{G}}(e)_{i-n_e} &  \mbox{otherwise}
	\end{cases}\]

@@

	\smallskip 
	Hence $(\id{E_{\mathcal{G}}}, \epsilon_{G})$ is a morphism $P(\sp(\mathcal{G}))\to \mathcal{G}$. Now, let $(f,g)\colon P(\mathcal{H})\to \mathcal{G}$ be a  morphism of pruned hypergraphs with $\mathcal{H}$ separated. Given a node $v$ in $V_{\mathcal{H}}$ we know that there exists exactly one $e_v$ such that $v$ occurs in $s_{\mathcal{H}}(e_v)$ or in $t_{\mathcal{H}}(e_v)$. Moreover, $v$ occurs exactly once and not in both. Let $i_v$ be the index at which it appears. We define
	\[\hat{g}(v):=\begin{cases}
		\iota_{f(e_v)}(\iota^{f(e_v)}_{i_v}(0)) & \text{if } v \text{ occurs in } s_{\mathcal{H}}(e)\\
		\iota_{f(e_v)}(\iota^{f(e_v)}_{n_e + i_v}(0)) & \text{if } v \text{ occurs in } t_{\mathcal{H}}(e)\\
	\end{cases}\]
	Computing we get
	\begin{gather*}
		\hat{g}^\star(s_{\mathcal{H}}(e))= \hat{g}^\star(\prod_{i=1}^{n_e} s_{\mathcal{H}}(e)_i)=\prod_{i=1}^{n_e}(\hat{g}(s_{\mathcal{H}}(e)_i))=\prod_{i=1}^{n_e}\iota_{f(e)}( \iota^{f(e)}_i(0))=s_{\sp(\mathcal{G})}(f(e))\\
		\hat{g}^\star(t_{\mathcal{H}}(e))= \hat{g}^\star(\prod_{i=1}^{m_e} t_{\mathcal{H}}(e)_i)=\prod_{i=1}^{m_e}(\hat{g}(t_{\mathcal{H}}(e)_i))=\prod_{i=1}^{m_e}\iota_{f(e)}( \iota^{f(e)}_{n_e+i}(0))=\prod_{i=n_e + 1}^{n_e + m_e}\iota_{f(e)}( \iota^{f(e)}_{i}(0))=t_{\sp(\mathcal{G})}(f(e))
	\end{gather*}
	proving that $(f, \hat{g})\colon \mathcal{H}\to\sp( \mathcal{G})$ is a morphism of $\pshyp$. Notice that	
	\begin{align*}
		&\epsilon_\mathcal{G}(\hat{g}(v))=\begin{cases}
			\epsilon_\mathcal{G}(\iota_{f(e_v)}(\iota^{f(e_v)}_{i_v}(0))) & \text{if } v \text{ occurs in } s_{\mathcal{H}}(e_v)\\
			\epsilon_\mathcal{G}(\iota_{f(e_v)}(\iota^{f(e_v)}_{n_e + i_v}(0))) & \text{if }v \text{ occurs in } t_{\mathcal{H}}(e_v)
		\end{cases}=	\begin{cases}
		\delta^{f(e_v)}_{i_v}(0)& \text{if }v \text{ occurs in } s_{\mathcal{G}}(e)\\
		\delta^{f(e_v)}_{i_v}(0)& \text{if } v \text{ occurs in } t_{\mathcal{G}}(e)
		\end{cases}\\ =&
		\begin{cases}
			s_{\mathcal{G}}(f(e_v))_{i_v}& \text{if } v \text{ occurs in } s_{\mathcal{G}}(e)\\
			t_{\mathcal{G}}(f(e_v))_{i_v}& \text{if } v \text{ occurs in } t_{\mathcal{G}}(e)
		\end{cases}= \begin{cases}
		(g^\star(s_{\mathcal{H}}(e_v)))_{i_v}& \text{if } v \text{ occurs in } s_{\mathcal{G}}(e)\\
		(g^\star(t_{\mathcal{H}}(e_v)))_{i_v}& \text{if } v \text{ occurs in } t_{\mathcal{G}}(e)\\
		\end{cases}\\=&\begin{cases}
		g(s_{\mathcal{H}}(e_v)_{i_v})& \text{if } v \text{ occurs in } s_{\mathcal{G}}(e)\\
		g(t_{\mathcal{H}}(e_v)_{i_v})& \text{if } v \text{ occurs in } t_{\mathcal{G}}(e)\\
		\end{cases}=g(v)
	\end{align*}
and so $(\id{E_{\mathcal{G}}}, \epsilon_{G})\circ P(f, \hat{g})=(f,g)$.

@@

\begin{proof}{\label{proof:pruned}}
	Let $\mathcal{G}$ be a pruned and separated hypergraph. We can define $F(\mathcal{G})$ as $(\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}}) \colon E_{\mathcal{G}}\to \mathbb{N}\times \mathbb{N}$. Given a morphism $(h,k)\colon \mathcal{G}\to \mathcal{H}$ and using \Cref{rem:le} we	have
	\[(\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}})\circ h = (\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}\circ h, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}}\circ h)=(\lgh_{V_{\mathcal{H}}}\circ k^\star\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{H}}}\circ k^\star\circ t_{\mathcal{G}})= (\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}})\]
	Thus we can define $F(h,k)$ as $h$ to get a functor $\pshyp\to \Set/(\mathbb{N}\times \mathbb{N})$. Let us show that it is an equivalence.
	\begin{itemize}
		\item $F$ is faithful. This follows at once from \Cref{rem:prun}.
		\item $F$ is full. Let $k\colon E_{\mathcal{G}}\to E_{\mathcal{H}}$ be an arrow such that 
		\[(\lgh_{V_{\mathcal{H}}}\circ s_{\mathcal{H}}, \lgh_{V_{\mathcal{H}}}\circ t_{\mathcal{H}})\circ f =(\lgh_{V_{\mathcal{G}}}\circ s_{\mathcal{G}}, \lgh_{V_{\mathcal{G}}}\circ t_{\mathcal{G}})\]

@@

		\noindent
		\begin{minipage}[l]{.78\linewidth}
		Since $\mathcal{G}$ is pruned and separated, $(V_{\mathcal{G}}, \{\iota^s_e\}_{e\in E_{\mathcal{G}}}+\{\iota^t_e\}_{e\in E_{\mathcal{G}}})$ is a coproduct. Given $v\in V^s_e$ and $w\in V^t_e$, let $i_v$ and $k_w$ be the unique indexes such that $\iota^s_e(v)=s_{\mathcal{G}}(e)_{i_v}$ and $\iota^t_e(w)= t_{\mathcal{G}}(e)_{k_w}$ and define
			\[\begin{split}
				g^s_e\colon V^s_{e} &\to V_{\mathcal{H}}\\
				v&\mapsto  s_{\mathcal{H}}(f(e))_{i_v}
			\end{split} \qquad \begin{split}
				g^t_e\colon V^t_{e} &\to V_{\mathcal{H}}\\
				w&\mapsto  t_{\mathcal{H}}(f(e))_{k_w}
			\end{split}\]
			so that there is an arrow $g\colon V_ {\mathcal{G}}\to \mathcal{H}$ fitting in the diagram aside.
		\end{minipage}\hfill
		\begin{minipage}[r]{.2\linewidth}
		\xymatrix@R=14pt@C=14pt{&V^s_{\mathcal{G}} \ar[dr]^{g^s_e} \ar@{>->}[dl]_{\iota^s_e}\\ V_{\mathcal{G}} \ar@{.>}[rr]^-{g} && V_{\mathcal{H}}}		
		\xymatrix@R=14pt@C=14pt{&V^t_{\mathcal{G}} \ar[dr]^{g^t_e} \ar@{>->}[dl]_{\iota^t_e}\\ V_{\mathcal{G}} \ar@{.>}[rr]^-{g} && V_{\mathcal{H}}}		
		\end{minipage}

		Given $e\in E_\mathcal{G}$, for every $1\leq i \leq \lgh_{V_{\mathcal{G}}}(s_{\mathcal{G}}(e))$ and $1\leq k \leq \lgh_{V_{\mathcal{G}}}(t_{\mathcal{G}}(e))$ we can compute to get
		\begin{align*}
		(g^\star(s_{\mathcal{G}}(e)))_i&=(g^\star(s_{\mathcal{G}}(e)))_i=g(s_{\mathcal{G}}(e)_i)=s_{\mathcal{H}}(f(e))_i\\
		(g^\star(t_{\mathcal{G}}(e)))_k&=(g^\star(t_{\mathcal{G}}(e)))_k=g(t_{\mathcal{G}}(e)_k)=t_{\mathcal{H}}(f(e))_k
		\end{align*}
